Practical Machine Learning - Week 1
============================================================
##### Identify patients who will be admitted to a hospital within the next year, using historical claims data [link](http://www.heritagehealthprize.com/hhp)  
[Kaggle](http://www.kaggle.com)

##### components of a predictor  
- question -> input data -> features -> algorithm -> parameters -> evaluation  

##### relative order of importance  
- question > data > feature > algorithm
- Often more data > better models

##### Properties of good features
- Lead to data compression
- Retain relevant information
- Are created based on expert application knowledge

##### Common mistakes
- Trying to automate feature selection
- Not paying attention to data-specific quirks
- Throwing away information unnecessarily

##### The "Best" Machine Learning Method
- Interpretable
- Simple
- Accurate
- Fast (to train and test)
- Scalable

##### Prediction is about accuracy tradeoffs
- Interpretability versus accuracy
- Speed versus accuracy
- Simplicity versus accuracy
- Scalability versus accuracy

##### SPAM Example
```{r}
library(kernlab)
data(spam)
head(spam)
```

##### Plot the density of "your"
```{r}
plot(density(spam$your[spam$type=="nonspam"]),
     col="blue",main="",xlab="Frequency of 'your'")
lines(density(spam$your[spam$type=="spam"]),col="red")
```

##### Find a value C, if frequency of 'yours'>C, then predict "spam"
```{r}
plot(density(spam$your[spam$type=="nonspam"]),
     col="blue",main="",xlab="Frequency of 'your'")
lines(density(spam$your[spam$type=="spam"]),col="red")
abline(v=0.5,col="black")
```
  
##### prediction based on 0.5 threshold
```{r}
prediction<-ifelse(spam$your>0.5,"spam","nonspam")
table(prediction,spam$type)/length(spam$type)
```

  
##### In sample and out of sample error
- In sample error: the error rate you get on the same data set you used to build your predictor. Resubstitution error
- Out of sample error: The eror rate you get on a new data set. Generalization error

```{r}
set.seed(333)
smallSpam = spam[sample(dim(spam)[1],size=10),]
spamLabel = (smallSpam$type=="spam")*1+1
plot(smallSpam$capitalAve,col=spamLabel)
```

##### Split data into
- Training: build the model
- Testing: validate the model (can not use multiple model on test set, that's training the model)
- Validation: Also use to validate the model (refine model on test set and validate on validation set)










