---
output: html_document
---
Practical Machine Learning - Week 2
============================================================
##### Caret Package
- preProcess
- Data splitting
    - createDataPartition
    - createResample
    - createTimeSlices
- Training/testing functions
- Confusion matrix

##### Machine learning algorithms: unify the prediction process, integrate most of the algorithms
```{r include=FALSE}
 library(caret)
 library(kernlab)
 library(ROCR)
```
##### Use createDataPartition to stratify the dataset into train and test
```{r}
 data(spam)
 inTrain = createDataPartition(y=spam$type,p=0.75,list=F)
 training = spam[inTrain,]
 testing = spam[-inTrain,]
 dim(training)
```
##### Fit a model
```{r include=FALSE}
set.seed(32423)
modelFit = train(type~.,data=training,method="glm")
```
```{r}
modelFit
modelFit$finalModel
```
##### Prediction
```{r}
predictions = predict(modelFit,newdata=testing,type="prob")
predictions = as.data.frame(predictions)
roc_pred = prediction(predictions$spam,testing$type)
roc_perf = performance(roc_pred,measure="tpr",x.measure="fpr")
plot(roc_perf,main="ROC Curve")
auc_perf = performance(roc_pred,measure="auc")
unlist(auc_perf@y.values)
```

##### Confusion matrix
```{r}
predictions_class = predict(modelFit,newdata=testing,type="raw")
confusionMatrix(predictions_class,testing$type)
```


# Data Slicing
##### Create k-fold
```{r}
set.seed(32323)
folds = createFolds(y=spam$type,k=10,list=T,returnTrain=T)
sapply(folds,length)
```

##### Resampling / bootstrapping
```{r}
set.seed(323233)
folds = createResample(y=spam$type,times=10,list=T)
sapply(folds,length)
```

##### Time Slices
```{r}
set.seed(3232333)
tme = 1:1000
folds = createTimeSlices(y=tme,initialWindow = 20, horizon = 10)
names(folds)
```

# Training Options
- preProcess
- weights
- metric
- trControl = trainControl
- tuneGrid
- tuneLength

#### Metric options
##### Continuous outcomes:
- RMSE (Root mean squared error)
- RSquare
##### Categorical outcomes:
- Accuracy
- Kappa

#### trainControl
- method: boot, boot632, cv, repeatedcv, LOOCV, etc.
- number


# Plotting Predictors
```{r include = F}
library(ISLR)
library(ggplot2)
library(caret)
```

##### Summary of Wage data
```{r}
data(Wage)
summary(Wage)
```

##### Build training set
```{r}
inTrain = createDataPartition(y=Wage$wage,p=0.7,list=F)
training = Wage[inTrain,]
testing = Wage[-inTrain,]
dim(training)
dim(testing)
```

##### feature plot
```{r}
featurePlot(x=training[,c("age","education","jobclass")],
            y=training$wage,
            plot="pairs")
```

##### Qplot
```{r}
qplot(age,wage,data=training)
qplot(age,wage,colour=jobclass,data=training)
```

##### Regression smoothers
```{r}
qq = qplot(age,wage,colour=education,data=training)
qq + geom_smooth(method="lm",formula=y~x)
```

##### cut2, making factors, binning and capping
```{r include=FALSE}
library(Hmisc)
cutWage = cut2(training$wage,g=3)
table(cutWage)
```

##### Boxplot
```{r}
p1 = qplot(cutWage,age,data=training,fill=cutWage,geom=c("boxplot"))
p1
t1 = table(cutWage,training$jobclass)
t1
prop.table(t1,1) # row %
```

##### Density plots
```{r}
qplot(wage,colour=education,data=training,geom="density")
```











